### 2020-09-02 Manuscript/pole_table.tex
flip the polarity of Ediacaran poles within the figures and table, add more detail to Ediacaran pole discussion

\begin{longtable}{p{1 in}p{1 in}rrrrrrrr}
\toprule
                       terrane &                                          unit name & rating &  site lon &  site lat &  plon &  plat &  A\$\_\{95\}\$ &                    age &                                     pole reference \\
\midrule
\endhead
\midrule
\multicolumn{10}{r}{{Continued on next page}} \\
\midrule
\endfoot

\bottomrule
\endlastfoot
             Laurentia-Wyoming &                            Stillwater Complex - C2 &      A &     249.2 &      45.2 & 335.8 & -83.6 &       4.0 &     2705\$\textasciicircum \{+4\}\$\$\_\{-4\}\$ &                                 \textbackslash cite\{Selkin2008a\} \\
  Baltica-Fennoscandia-Karelia &                                Koitere sanukitoids &      B &      30.6 &      63.3 & 192.5 & -67.5 &      19.5 &     2684\$\textasciicircum \{+2\}\$\$\_\{-2\}\$ &                                                NaN \\
      Laurentia-Superior(East) &                       Otto Stock dikes and aureole &      B &     279.9 &      48.0 & 227.0 &  69.0 &       4.8 &     2676\$\textasciicircum \{+5\}\$\$\_\{-5\}\$ &                               \textbackslash cite\{Pullaiah1975b\} \\
               Laurentia-Slave &                                       Defeat Suite &      B &     245.5 &      62.5 &  64.0 &  -1.0 &      15.0 &     2625\$\textasciicircum \{+5\}\$\$\_\{-5\}\$ &                               \textbackslash cite\{Mitchell2014a\} \\
          Baltica-Fennoscandia &                   Shalskiy thick gabbronorite dike &      B &      36.0 &      61.9 & 222.1 &  22.7 &      11.5 &     2511\$\textasciicircum \{+2\}\$\$\_\{-2\}\$ &                                                NaN \\
      Laurentia-Superior(East) &                         Ptarmigan-Mistassini dikes &      B &     287.0 &      54.0 & 213.0 & -45.3 &      13.8 &     2505\$\textasciicircum \{+2\}\$\$\_\{-2\}\$ &                                  \textbackslash cite\{Evans2010a\} \\
          Baltica-Fennoscandia &                     Generalskaya Layered Intrusion &      B &      31.0 &      69.4 & 292.7 & -42.5 &      10.4 &     2505\$\textasciicircum \{+2\}\$\$\_\{-2\}\$ &                                                NaN \\
          Baltica-Fennoscandia &                              Monchegorsk Intrusion &      B &      33.0 &      68.0 & 265.3 &   1.3 &       9.9 &     2504\$\textasciicircum \{+2\}\$\$\_\{-2\}\$ &                                                NaN \\
          Baltica-Fennoscandia &  Avdeev gabbronorite and thin Shalskiy diabase ... &      B &      36.0 &      61.9 & 243.5 & -12.3 &      14.0 &   2476\$\textasciicircum \{+35\}\$\$\_\{-35\}\$ &                                                NaN \\
      Laurentia-Superior(East) &                                 Matachewan dikes R &      A &     278.0 &      48.0 & 238.3 & -44.1 &       1.6 &   2466\$\textasciicircum \{+23\}\$\$\_\{-23\}\$ &                                  \textbackslash cite\{Evans2010a\} \\
          Baltica-Fennoscandia &                          Imandra Layered Intrusion &      B &      33.1 &      67.6 & 280.3 & -16.1 &       7.8 &   2446\$\textasciicircum \{+39\}\$\$\_\{-39\}\$ &                                                NaN \\
          Baltica-Fennoscandia &                                     Karelian dikes &      B &      30.7 &      66.0 & 278.7 & -19.9 &       6.1 &     2446\$\textasciicircum \{+5\}\$\$\_\{-5\}\$ &                                                NaN \\
      Laurentia-Superior(East) &                                 Matachewan dikes N &      A &     278.0 &      48.0 & 239.5 & -52.3 &       2.4 &     2446\$\textasciicircum \{+3\}\$\$\_\{-3\}\$ &                                  \textbackslash cite\{Evans2010a\} \\
          Baltica-Fennoscandia &                            Taivalkoski 2.3 Ga dike &     C+ &      28.1 &      65.5 & 257.3 &  20.4 &       6.3 &   2339\$\textasciicircum \{+18\}\$\$\_\{-18\}\$ &                                                NaN \\
               Laurentia-Slave &                                       Malley dikes &      A &     249.8 &      64.2 & 310.0 & -50.8 &       6.7 &     2231\$\textasciicircum \{+2\}\$\$\_\{-2\}\$ &                                 \textbackslash cite\{Buchan2012a\} \\
      Laurentia-Superior(East) &                                   Senneterre dikes &      A &     283.0 &      49.0 & 284.3 & -15.3 &       5.5 &     2218\$\textasciicircum \{+6\}\$\$\_\{-6\}\$ &                                 \textbackslash cite\{Buchan1993a\} \\
      Laurentia-Superior(East) &                                 Nipissing N1 sills &      A &     279.0 &      47.0 & 272.0 & -17.0 &      10.0 &     2217\$\textasciicircum \{+4\}\$\$\_\{-4\}\$ &                                 \textbackslash cite\{Buchan2000a\} \\
               Laurentia-Slave &                                       Dogrib dikes &      A &     245.5 &      62.5 & 315.0 & -31.0 &       7.0 &     2193\$\textasciicircum \{+2\}\$\$\_\{-2\}\$ &                               \textbackslash cite\{Mitchell2014a\} \\
      Laurentia-Superior(East) &                                  Biscotasing dikes &      A &     280.0 &      48.0 & 223.9 &  26.0 &       7.0 &     2170\$\textasciicircum \{+3\}\$\$\_\{-3\}\$ &                                  \textbackslash cite\{Evans2010a\} \\
             Laurentia-Wyoming &    Rabbit Creek, Powder River and South Path dikes &      A &     252.8 &      43.9 & 339.2 &  65.5 &       7.6 &    2160\$\textasciicircum \{+11\}\$\$\_\{-8\}\$ &                                 \textbackslash cite\{Kilian2015a\} \\
               Laurentia-Slave &                                        Indin dikes &      A &     245.6 &      62.5 & 256.0 & -36.0 &       7.0 &    2126\$\textasciicircum \{+3\}\$\$\_\{-18\}\$ &                                 \textbackslash cite\{Buchan2016a\} \\
      Laurentia-Superior(West) &                                   Marathon dikes N &      A &     275.0 &      49.0 & 198.2 &  45.4 &       7.7 &     2124\$\textasciicircum \{+3\}\$\$\_\{-3\}\$ &                                  \textbackslash cite\{Halls2008a\} \\
      Laurentia-Superior(West) &                                   Marathon dikes R &      A &     275.0 &      49.0 & 182.2 &  55.1 &       7.5 &     2104\$\textasciicircum \{+3\}\$\$\_\{-3\}\$ &                                  \textbackslash cite\{Halls2008a\} \\
      Laurentia-Superior(West) &                                 Cauchon Lake dikes &      A &     263.0 &      56.0 & 180.9 &  53.8 &       7.7 &     2091\$\textasciicircum \{+2\}\$\$\_\{-2\}\$ &                                  \textbackslash cite\{Evans2010a\} \\
      Laurentia-Superior(West) &                                 Fort Frances dikes &      A &     266.0 &      48.0 & 184.6 &  42.8 &       6.1 &     2077\$\textasciicircum \{+5\}\$\$\_\{-5\}\$ &                                  \textbackslash cite\{Evans2010a\} \\
      Laurentia-Superior(East) &                                   Lac Esprit dikes &      A &     282.0 &      53.0 & 170.5 &  62.0 &       6.4 &     2069\$\textasciicircum \{+1\}\$\$\_\{-1\}\$ &                                  \textbackslash cite\{Evans2010a\} \\
              Baltica-Sarmatia &               Volyn-Dniestr-Bug intrusions A-group &      B &      29.8 &      49.6 & 182.9 &  15.7 &      13.7 &   2061\$\textasciicircum \{+20\}\$\$\_\{-20\}\$ &                                                NaN \\
          Baltica-Fennoscandia &                               Kuetsyarvi Formation &      B &      29.5 &      69.5 & 300.8 &  24.7 &      16.4 &     2060\$\textasciicircum \{+8\}\$\$\_\{-8\}\$ &                                                NaN \\
      Laurentia-Greenland-Nain &                                    Kangamiut dikes &      B &     307.0 &      66.0 & 273.8 &  17.1 &       2.7 &   2042\$\textasciicircum \{+12\}\$\$\_\{-12\}\$ &                                 \textbackslash cite\{Fahrig1976b\} \\
               Laurentia-Slave &                                  Lac de Gras dikes &      A &     249.6 &      64.4 & 267.9 &  11.8 &       7.1 &     2026\$\textasciicircum \{+5\}\$\$\_\{-5\}\$ &                                 \textbackslash cite\{Buchan2009a\} \\
              Baltica-Sarmatia &                       Volyn-Dniestr-Bug intrusions &      B &      27.4 &      50.9 & 140.4 &  64.4 &      12.7 &   2000\$\textasciicircum \{+10\}\$\$\_\{-10\}\$ &                                                NaN \\
      Laurentia-Superior(East) &                                        Minto dikes &      A &     285.0 &      57.0 & 171.5 &  38.7 &      13.1 &     1998\$\textasciicircum \{+2\}\$\$\_\{-2\}\$ &                                  \textbackslash cite\{Evans2010a\} \\
             Baltica - Karelia &                           Onego 1975 Ma Intrusions &      A &      34.0 &      62.1 & 101.5 &  44.4 &       6.3 &     1976\$\textasciicircum \{+9\}\$\$\_\{-9\}\$ &                                                NaN \\
          Baltica-Fennoscandia &                                    Konchozero Sill &      B &      34.0 &      62.1 & 281.9 & -14.2 &      10.4 &   1974\$\textasciicircum \{+27\}\$\$\_\{-27\}\$ &                                                NaN \\
               Laurentia-Slave &                                    Rifle Formation &      B &     252.9 &      65.9 & 341.0 &  14.0 &       7.7 &     1963\$\textasciicircum \{+6\}\$\$\_\{-6\}\$ &                                  \textbackslash cite\{Evans1981a\} \\
          Baltica-Fennoscandia &             Tsuomasavarri Gabbro-Diorite Intrusion &      B &      28.3 &      69.9 & 247.3 &  40.2 &       6.0 &     1931\$\textasciicircum \{+2\}\$\$\_\{-2\}\$ &                                                NaN \\
                 Laurentia-Rae &                             Clearwater Anorthosite &      B &     251.6 &      57.1 & 311.8 &   6.5 &       2.9 &     1917\$\textasciicircum \{+7\}\$\$\_\{-7\}\$ &                                  \textbackslash cite\{Halls1999a\} \\
             Laurentia-Wyoming &                         Sourdough mafic dike swarm &      A &    -108.3 &      44.7 & 292.0 &  49.2 &       8.1 &     1899\$\textasciicircum \{+5\}\$\$\_\{-5\}\$ &                                 \textbackslash cite\{Kilian2016b\} \\
               Laurentia-Slave &                                   Ghost Dike Swarm &      A &     244.6 &      62.6 & 286.0 &  -2.0 &       6.0 &     1887\$\textasciicircum \{+5\}\$\$\_\{-9\}\$ &                                 \textbackslash cite\{Buchan2016a\} \\
          Baltica-Fennoscandia &                         Kiuruvesi-Pielavesi gabbro &      B &      26.8 &      63.2 & 231.3 &  41.1 &       5.4 &     1886\$\textasciicircum \{+6\}\$\$\_\{-5\}\$ &                                                NaN \\
               Laurentia-Slave &                           Mean Seton/Akaitcho/Mara &      B &     250.0 &      65.0 & 260.0 &  -6.0 &       4.0 &     1885\$\textasciicircum \{+5\}\$\$\_\{-5\}\$ &                               \textbackslash cite\{Mitchell2010c\} \\
               Laurentia-Slave &                     Mean Kahochella, Peacock Hills &      B &     250.0 &      65.0 & 285.0 & -12.0 &       7.0 &     1882\$\textasciicircum \{+4\}\$\$\_\{-4\}\$ &                               \textbackslash cite\{Mitchell2010c\} \\
      Laurentia-Superior(West) &                                Molson (B+C2) dikes &      A &     262.0 &      55.0 & 218.0 &  28.9 &       3.8 &     1879\$\textasciicircum \{+6\}\$\$\_\{-6\}\$ &                                  \textbackslash cite\{Evans2010a\} \\
               Laurentia-Slave &                                 Takiyuak Formation &      B &     246.9 &      66.1 & 249.0 & -13.0 &       8.0 &   1876\$\textasciicircum \{+10\}\$\$\_\{-10\}\$ &                                 \textbackslash cite\{Irving1979a\} \\
               Laurentia-Slave &          Douglas Peninsula Formation, Pethei Group &      B &     249.7 &      62.8 & 258.0 & -18.0 &      14.2 &   1876\$\textasciicircum \{+10\}\$\$\_\{-10\}\$ &                                 \textbackslash cite\{Irving1979a\} \\
               Laurentia-Slave &               Pearson A/Peninsular/Kilohigok sills &      A &     250.0 &      65.0 & 269.0 & -22.0 &       6.0 &     1870\$\textasciicircum \{+4\}\$\$\_\{-4\}\$ &                               \textbackslash cite\{Mitchell2010c\} \\
            Laurentia-Superior &                         Haig/Flaherty/Sutton Mean  &      B &     281.0 &      56.2 & 245.8 &   1.0 &       3.9 &     1870\$\textasciicircum \{+1\}\$\$\_\{-1\}\$ &  Nordic workshop calculation based on data of \textbackslash ... \\
        Baltica - Fennoscandia &                                      Keuruu dikes  &     B  &      24.7 &      62.3 & 230.9 &  45.7 &       5.5 &   1869\$\textasciicircum \{+10\}\$\$\_\{-10\}\$ &                                                NaN \\
 Laurentia-Trans-Hudson orogen &                                Boot-Phantom Pluton &      B &     258.1 &      54.7 & 275.4 &  62.4 &       7.9 &     1838\$\textasciicircum \{+1\}\$\$\_\{-1\}\$ &                                 \textbackslash cite\{Symons1999a\} \\
                 Laurentia-Rae &                                      Sparrow dikes &      B &     250.2 &      61.6 & 291.0 &  12.0 &       7.9 &     1827\$\textasciicircum \{+4\}\$\$\_\{-4\}\$ &                                \textbackslash cite\{McGlynn1974a\} \\
                 Laurentia-Rae &                                   Martin Formation &      A &     251.4 &      59.6 & 288.0 &  -9.0 &       8.5 &     1818\$\textasciicircum \{+4\}\$\$\_\{-4\}\$ &                                  \textbackslash cite\{Evans1973a\} \\
          Baltica-Fennoscandia &                                    Notträsk gabbro &      B &      21.8 &      65.8 & 216.2 &  43.5 &       6.1 &     1806\$\textasciicircum \{+6\}\$\$\_\{-6\}\$ &                                                NaN \\
          Baltica-Fennoscandia &                                      Kallax gabbro &      B &      22.0 &      65.5 & 209.0 &  49.0 &       nan &     1800\$\textasciicircum \{+5\}\$\$\_\{-6\}\$ &                                                NaN \\
          Baltica-Fennoscandia &                 Lake Ladoga, Mean intr. \& dikes, A &      B &      30.0 &      61.3 & 229.1 &  50.9 &       nan &   1800\$\textasciicircum \{+19\}\$\$\_\{-56\}\$ &                                                NaN \\
          Baltica-Fennoscandia &                                      Hoting gabbro &      A &      16.2 &      64.2 & 233.3 &  43.0 &      10.9 &   1786\$\textasciicircum \{+10\}\$\$\_\{-10\}\$ &                                                NaN \\
                     Laurentia &                                      Dubawnt Group &      B &     265.6 &      64.1 & 277.0 &   7.0 &       8.0 &   1785\$\textasciicircum \{+35\}\$\$\_\{-35\}\$ &                                   \textbackslash cite\{Park1973a\} \\
          Baltica-Fennoscandia &                              Småland intrusives -C &      A &      15.7 &      57.1 & 182.7 &  45.7 &       8.0 &     1777\$\textasciicircum \{+8\}\$\$\_\{-8\}\$ &                                                NaN \\
          Baltica-Fennoscandia &                                 Shoksha Sandstones &      B &      33.8 &      61.3 & 221.1 &  39.7 &       4.0 &   1775\$\textasciicircum \{+25\}\$\$\_\{-25\}\$ &                                                NaN \\
 Laurentia-Trans-Hudson orogen &                            Deschambault Pegmatites &      B &     256.7 &      54.9 & 276.0 &  67.5 &       7.7 &     1766\$\textasciicircum \{+5\}\$\$\_\{-5\}\$ &                                 \textbackslash cite\{Symons2000a\} \\
 Laurentia-Trans-Hudson orogen &                                   Jan Lake Granite &      B &     257.2 &      54.9 & 264.3 &  24.3 &      16.9 &     1758\$\textasciicircum \{+1\}\$\$\_\{-1\}\$ &                                   \textbackslash cite\{Gala1995a\} \\
              Baltica-Sarmatia &                       Volyn-Dniestr-Bug intrusions &      A &      28.6 &      50.6 & 169.1 &  26.5 &       3.9 &   1755\$\textasciicircum \{+15\}\$\$\_\{-15\}\$ &                                                NaN \\
          Baltica-Fennoscandia &                                MEAN Ropruchey sill &      B &      35.3 &      61.4 & 217.0 &  39.1 &       8.6 &     1751\$\textasciicircum \{+3\}\$\$\_\{-3\}\$ &                                                NaN \\
                     Laurentia &                                      Cleaver dikes &      A &     242.0 &      67.5 & 276.7 &  19.4 &       6.1 &     1741\$\textasciicircum \{+5\}\$\$\_\{-5\}\$ &                                 \textbackslash cite\{Irving2004a\} \\
              Baltica-Sarmatia &                       Volyn-Dniestr-Bug intrusions &      B &      29.6 &      50.0 & 163.2 &  10.7 &      10.2 &   1722\$\textasciicircum \{+12\}\$\$\_\{-12\}\$ &                                                NaN \\
        Baltica - Fennoscandia &                                   Häme dike swarm  &      B &      25.1 &      61.4 & 209.8 &  23.6 &      14.7 &   1647\$\textasciicircum \{+14\}\$\$\_\{-14\}\$ &                                                NaN \\
                       Baltica &                        Sipoo Quartz Porphyry dikes &      B &      25.3 &      60.3 & 180.4 &  26.4 &       9.4 &   1633\$\textasciicircum \{+10\}\$\$\_\{-10\}\$ &                                                NaN \\
           Laurentia-Greenland &                        Melville Bugt diabase dikes &      B &     303.0 &      74.6 & 273.8 &   5.0 &       8.7 &     1633\$\textasciicircum \{+5\}\$\$\_\{-5\}\$ &                                  \textbackslash cite\{Halls2011a\} \\
                       Baltica &                              Quartz porphyry dikes &      B &      26.8 &      61.3 & 175.4 &  30.2 &       9.4 &   1631\$\textasciicircum \{+10\}\$\$\_\{-10\}\$ &                                                NaN \\
                       Baltica &                                 Hoting basic dikes &      B &      16.2 &      64.2 & 146.7 &  21.9 &      13.7 &   1614\$\textasciicircum \{+24\}\$\$\_\{-24\}\$ &                                                NaN \\
                     Laurentia &                            Western Channel Diabase &      A &     242.2 &      66.4 & 245.0 &   9.0 &       6.6 &     1590\$\textasciicircum \{+3\}\$\$\_\{-3\}\$ &                                 \textbackslash cite\{Irving1972a\} \\
                       Baltica &                                   Åland dike swarm &      A &      20.0 &      60.0 & 191.4 &  23.7 &       2.8 &   1578\$\textasciicircum \{+12\}\$\$\_\{-12\}\$ &                                                NaN \\
                       Baltica &                            Satakunta dike swarm -C &      A &      21.5 &      62.0 & 188.1 &  29.3 &       6.6 &   1578\$\textasciicircum \{+12\}\$\$\_\{-12\}\$ &                                                NaN \\
                       Baltica &                                  Ragunda Formation &      B &      16.1 &      63.3 & 166.6 &  51.6 &       7.1 &   1506\$\textasciicircum \{+13\}\$\$\_\{-13\}\$ &                                                NaN \\
                     Laurentia &                 St.Francois Mountains Acidic Rocks &      A &     269.5 &      37.5 & 219.0 & -13.2 &       6.1 &   1476\$\textasciicircum \{+16\}\$\$\_\{-16\}\$ &                                  \textbackslash cite\{Meert2002b\} \\
                       Baltica &                           MEAN Bunkris/Glysjon/Oje &      B &      13.5 &      61.5 & 179.8 &  28.3 &      13.2 &     1469\$\textasciicircum \{+9\}\$\$\_\{-9\}\$ &                                                NaN \\
                     Laurentia &                               Michikamau Intrusion &      A &     296.0 &      54.5 & 217.5 &  -1.5 &       4.7 &     1460\$\textasciicircum \{+5\}\$\$\_\{-5\}\$ &                                 \textbackslash cite\{Emslie1976a\} \\
                     Laurentia &                                  Spokane Formation &      A &     246.8 &      48.2 & 215.5 & -24.8 &       4.7 &   1458\$\textasciicircum \{+13\}\$\$\_\{-13\}\$ &                                 \textbackslash cite\{Elston2002a\} \\
                       Baltica &                    Lake Ladoga basalt, sill, dikes &      A &      31.2 &      61.5 & 177.1 &  15.2 &       5.5 &   1452\$\textasciicircum \{+12\}\$\$\_\{-12\}\$ &                                                NaN \\
                     Laurentia &                                 Snowslip Formation &      A &     245.9 &      47.9 & 210.2 & -24.9 &       3.5 &   1450\$\textasciicircum \{+14\}\$\$\_\{-14\}\$ &                                 \textbackslash cite\{Elston2002a\} \\
                     Laurentia &                                 Tobacco Root dikes &      B &     247.6 &      47.4 & 216.1 &   8.7 &      10.5 &   1448\$\textasciicircum \{+49\}\$\$\_\{-49\}\$ &                                 \textbackslash cite\{Harlan2008a\} \\
                     Laurentia &                                       Purcell Lava &      A &     245.1 &      49.4 & 215.6 & -23.6 &       4.8 &     1443\$\textasciicircum \{+7\}\$\$\_\{-7\}\$ &                                 \textbackslash cite\{Elston2002a\} \\
                     Laurentia &                          Rocky Mountain intrusions &      B &     253.8 &      40.3 & 217.4 & -11.9 &       9.7 &   1430\$\textasciicircum \{+15\}\$\$\_\{-15\}\$ &  Nordic workshop calculation based on data of \textbackslash ... \\
                     Laurentia &                                   Mistastin Pluton &      B &     296.3 &      55.6 & 201.5 &  -1.0 &       7.6 &   1425\$\textasciicircum \{+25\}\$\$\_\{-25\}\$ &                                 \textbackslash cite\{Fahrig1976a\} \\
                     Laurentia &                                 McNamara Formation &      A &     246.4 &      46.9 & 208.3 & -13.5 &       6.7 &     1401\$\textasciicircum \{+6\}\$\$\_\{-6\}\$ &                                 \textbackslash cite\{Elston2002a\} \\
                     Laurentia &         Pilcher, Garnet Range and Libby Formations &      A &     246.4 &      46.7 & 215.3 & -19.2 &       7.7 &   1385\$\textasciicircum \{+23\}\$\$\_\{-23\}\$ &                                 \textbackslash cite\{Elston2002a\} \\
           Laurentia-Greenland &                                Zig-Zag Dal Basalts &      B &     334.8 &      81.2 & 242.8 &  12.0 &       3.8 &     1382\$\textasciicircum \{+2\}\$\$\_\{-2\}\$ &                              \textbackslash cite\{Marcussen1983a\} \\
           Laurentia-Greenland &                      Victoria Fjord dolerite dikes &      B &     315.3 &      81.5 & 231.7 &  10.3 &       4.3 &     1382\$\textasciicircum \{+2\}\$\$\_\{-2\}\$ &                             \textbackslash cite\{Abrahamsen1987a\} \\
           Laurentia-Greenland &                              Midsommersoe Dolerite &      B &     333.4 &      81.6 & 242.0 &   6.9 &       5.1 &     1382\$\textasciicircum \{+2\}\$\$\_\{-2\}\$ &                              \textbackslash cite\{Marcussen1983a\} \\
                       Baltica &                                             Mashak &      B &      57.0 &      54.0 & 193.0 &   1.8 &      14.8 &   1376\$\textasciicircum \{+10\}\$\$\_\{-10\}\$ &                                                NaN \\
                     Laurentia &                                   Nain Anorthosite &      B &     298.2 &      56.5 & 206.7 &  11.7 &       2.2 &   1305\$\textasciicircum \{+15\}\$\$\_\{-15\}\$ &                                 \textbackslash cite\{Murthy1978a\} \\
           Laurentia-Greenland &                             North Qoroq intrusives &      B &     314.6 &      61.1 & 202.6 &  13.2 &       8.3 &     1275\$\textasciicircum \{+1\}\$\$\_\{-1\}\$ &                                  \textbackslash cite\{Piper1992a\} \\
           Laurentia-Greenland &                                  Kungnat Ring dike &      B &     311.7 &      61.2 & 198.7 &   3.4 &       3.2 &     1275\$\textasciicircum \{+2\}\$\$\_\{-2\}\$ &                                  \textbackslash cite\{Piper1977b\} \\
                     Laurentia &                         Mackenzie dikes grand mean &      A &     250.0 &      65.0 & 190.0 &   4.0 &       5.0 &     1267\$\textasciicircum \{+2\}\$\$\_\{-2\}\$ &                                 \textbackslash cite\{Buchan2000a\} \\
                       Baltica &                       MEAN post-Jotnian intrusions &      A &      18.1 &      62.4 & 159.1 &  -1.8 &       3.4 &   1258\$\textasciicircum \{+12\}\$\$\_\{-12\}\$ &                                                NaN \\
           Laurentia-Greenland &                         West Gardar Dolerite dikes &      B &     311.7 &      61.2 & 201.7 &   8.7 &       6.6 &     1244\$\textasciicircum \{+8\}\$\$\_\{-8\}\$ &                                  \textbackslash cite\{Piper1977b\} \\
           Laurentia-Greenland &                      West Gardar Lamprophyre dikes &      B &     311.7 &      61.2 & 206.4 &   3.2 &       7.2 &   1238\$\textasciicircum \{+11\}\$\$\_\{-11\}\$ &                                  \textbackslash cite\{Piper1977b\} \\
                     Laurentia &                             Sudbury dikes Combined &      A &     278.6 &      46.3 & 192.8 &  -2.5 &       2.5 &     1237\$\textasciicircum \{+5\}\$\$\_\{-5\}\$ &                                 \textbackslash cite\{Palmer1977a\} \\
            Laurentia-Scotland &                                        Stoer Group &      B &     354.5 &      58.0 & 238.4 &  37.2 &       7.7 &   1199\$\textasciicircum \{+70\}\$\$\_\{-70\}\$ &                        Nordic workshop calculation \\
           Laurentia-Greenland &                                 Hviddal Giant dike &      B &     313.7 &      60.9 & 215.3 &  33.2 &       9.6 &     1184\$\textasciicircum \{+5\}\$\$\_\{-5\}\$ &                                  \textbackslash cite\{Piper1977a\} \\
           Laurentia-Greenland &                                     Narssaq Gabbro &      B &     313.8 &      60.9 & 225.4 &  31.6 &       9.7 &     1184\$\textasciicircum \{+5\}\$\$\_\{-5\}\$ &                                  \textbackslash cite\{Piper1977a\} \\
           Laurentia-Greenland &                                  South Qoroq Intr. &      A &     314.6 &      61.1 & 215.9 &  41.8 &      13.1 &     1163\$\textasciicircum \{+2\}\$\$\_\{-2\}\$ &                                  \textbackslash cite\{Piper1992a\} \\
           Laurentia-Greenland &                                 Giant Gabbro dikes &      B &     313.7 &      60.9 & 226.1 &  42.3 &       9.4 &     1163\$\textasciicircum \{+2\}\$\$\_\{-2\}\$ &                                  \textbackslash cite\{Piper1977a\} \\
           Laurentia-Greenland &                               NE-SW Trending dikes &      B &     314.6 &      61.1 & 230.8 &  33.4 &       5.7 &     1160\$\textasciicircum \{+5\}\$\$\_\{-5\}\$ &                                  \textbackslash cite\{Piper1992a\} \\
                     Laurentia &                          Ontario lamprophyre dikes &     NR &     273.3 &      48.8 & 223.3 &  58.0 &       9.2 &   1143\$\textasciicircum \{+12\}\$\$\_\{-12\}\$ &                                 \textbackslash cite\{Piispa2018a\} \\
                     Laurentia &                                      Abitibi dikes &      A &     279.0 &      48.0 & 215.5 &  48.8 &      14.1 &     1141\$\textasciicircum \{+2\}\$\$\_\{-2\}\$ &                                  \textbackslash cite\{Ernst1993a\} \\
                       Baltica &                        Salla dike (width 60-100 m) &      B &      28.8 &      66.8 & 111.6 &  70.6 &       8.1 &     1122\$\textasciicircum \{+5\}\$\$\_\{-3\}\$ &                                                NaN \\
                     Laurentia &                            Nipigon sills and lavas &      A &     270.9 &      49.1 & 217.8 &  47.2 &       4.0 &     1109\$\textasciicircum \{+2\}\$\$\_\{-2\}\$ &  Nordic workshop calculation based on data of \textbackslash ... \\
                     Laurentia &             Lowermost Mamainse Point volcanics -R1 &      A &     275.3 &      47.1 & 227.0 &  49.5 &       5.3 &     1109\$\textasciicircum \{+2\}\$\$\_\{-3\}\$ &                         \textbackslash cite\{Swanson-Hysell2014a\} \\
                     Laurentia &                           Lower Osler volcanics -R &      A &     272.3 &      48.8 & 218.6 &  40.9 &       4.8 &     1108\$\textasciicircum \{+3\}\$\$\_\{-3\}\$ &                         \textbackslash cite\{Swanson-Hysell2014b\} \\
                     Laurentia &                          Middle Osler volcanics -R &      A &     272.4 &      48.8 & 211.3 &  42.7 &       8.2 &     1107\$\textasciicircum \{+4\}\$\$\_\{-4\}\$ &                         \textbackslash cite\{Swanson-Hysell2014b\} \\
                     Laurentia &                           Upper Osler volcanics -R &      A &     272.4 &      48.7 & 203.4 &  42.3 &       3.7 &     1105\$\textasciicircum \{+1\}\$\$\_\{-1\}\$ &  \textbackslash cite\{Halls1974a, Swanson-Hysell2014b, Swanson... \\
                     Laurentia &                 Lower Mamainse Point volcanics -R2 &      A &     275.3 &      47.1 & 205.2 &  37.5 &       4.5 &     1105\$\textasciicircum \{+3\}\$\$\_\{-4\}\$ &                         \textbackslash cite\{Swanson-Hysell2014a\} \\
                     Laurentia &     Mamainse Point volcanics -C (lower N, upper R) &      A &     275.3 &      47.1 & 189.7 &  36.1 &       4.9 &     1101\$\textasciicircum \{+1\}\$\$\_\{-1\}\$ &                         \textbackslash cite\{Swanson-Hysell2014a\} \\
                     Laurentia &                               North Shore lavas -N &      A &     268.7 &      46.3 & 181.7 &  31.1 &       2.1 &     1097\$\textasciicircum \{+3\}\$\$\_\{-3\}\$ &              \textbackslash cite\{Tauxe2009a,Swanson-Hysell2019a\} \\
                     Laurentia &                              Chengwatana Volcanics &      B &     267.3 &      45.4 & 186.1 &  30.9 &       8.2 &     1095\$\textasciicircum \{+2\}\$\$\_\{-2\}\$ &                                   \textbackslash cite\{Kean1997a\} \\
                     Laurentia &                             Portage Lake Volcanics &      A &     271.2 &      47.0 & 182.5 &  27.5 &       2.3 &     1095\$\textasciicircum \{+3\}\$\$\_\{-3\}\$ &  \textbackslash cite\{Books1972a, Hnat2006a\} as calculated in ... \\
                     Laurentia &              Uppermost Mamainse Point volcanics -N &      A &     275.3 &      47.1 & 183.2 &  31.2 &       2.5 &     1094\$\textasciicircum \{+6\}\$\$\_\{-4\}\$ &                         \textbackslash cite\{Swanson-Hysell2014a\} \\
                     Laurentia &                    Cardenas Basalts and Intrusions &      B &     248.1 &      36.1 & 185.0 &  32.0 &       8.0 &     1091\$\textasciicircum \{+5\}\$\$\_\{-5\}\$ &                                   \textbackslash cite\{Weil2003a\} \\
                     Laurentia &                           Schroeder Lutsen Basalts &      A &     269.1 &      47.5 & 187.8 &  27.1 &       3.0 &     1090\$\textasciicircum \{+2\}\$\$\_\{-7\}\$ &                              \textbackslash cite\{Fairchild2017a\} \\
                     Laurentia &                        Central Arizona diabases -N &      A &     249.2 &      33.7 & 175.3 &  15.7 &       7.0 &   1088\$\textasciicircum \{+11\}\$\$\_\{-11\}\$ &                               \textbackslash cite\{Donadini2011b\} \\
                     Laurentia &                                   Lake Shore Traps &      A &     271.9 &      47.6 & 186.4 &  23.1 &       4.0 &     1086\$\textasciicircum \{+1\}\$\$\_\{-1\}\$ &                                \textbackslash cite\{Kulakov2013a\} \\
                     Laurentia &                      Michipicoten Island Formation &      A &     274.3 &      47.7 & 174.7 &  17.0 &       4.4 &     1084\$\textasciicircum \{+1\}\$\$\_\{-1\}\$ &                              \textbackslash cite\{Fairchild2017a\} \\
                     Laurentia &                                     Nonesuch Shale &      B &     271.5 &      47.0 & 178.1 &   7.6 &       5.5 &    1080\$\textasciicircum \{+4\}\$\$\_\{-10\}\$ &                                  \textbackslash cite\{Henry1977a\} \\
                     Laurentia &                                    Freda Sandstone &      B &     271.5 &      47.0 & 179.0 &   2.2 &       4.2 &   1070\$\textasciicircum \{+14\}\$\$\_\{-10\}\$ &                                  \textbackslash cite\{Henry1977a\} \\
                       Baltica &                             MEAN Bamble Intrusions &     C+ &       9.5 &      58.9 & 222.5 & -15.1 &      19.3 &   1048\$\textasciicircum \{+73\}\$\$\_\{-73\}\$ &                                                NaN \\
                       Baltica &                           Laanila-Ristijarvi dikes &      B &      28.1 &      68.7 & 212.2 &  -2.1 &      13.8 &   1044\$\textasciicircum \{+52\}\$\$\_\{-52\}\$ &                                                NaN \\
                     Laurentia &                              Haliburton Intrusions &      B &     281.4 &      45.0 & 141.9 & -32.6 &       6.3 &   1015\$\textasciicircum \{+15\}\$\$\_\{-15\}\$ &                                \textbackslash cite\{Warnock2000a\} \\
                     Laurentia &                        Adirondack fayalite granite &     NR &     285.5 &      44.0 & 132.7 & -28.4 &       6.9 &    990\$\textasciicircum \{+20\}\$\$\_\{-20\}\$ &                                  \textbackslash cite\{Brown2012a\} \\
                     Laurentia &               Adirondack metamorphic\textbackslash nanorthosites &     NR &     286.0 &      44.0 & 149.0 & -25.1 &      11.6 &    970\$\textasciicircum \{+20\}\$\$\_\{-20\}\$ &                                  \textbackslash cite\{Brown2012a\} \\
                     Laurentia &                       Adirondack Microcline gneiss &     NR &     285.0 &      44.0 & 151.1 & -18.4 &      10.5 &    960\$\textasciicircum \{+20\}\$\$\_\{-20\}\$ &                                  \textbackslash cite\{Brown2012a\} \\
                       Baltica &                            MEAN Blekinga dolerites &      B &      15.0 &      56.1 & 247.0 &  13.0 &      16.0 &      950\$\textasciicircum \{+4\}\$\$\_\{-4\}\$ &                                                NaN \\
                       Baltica &               dikes of central and southern Sweden &      A &      14.2 &      58.4 & 240.7 &  -0.9 &       6.7 &    945\$\textasciicircum \{+10\}\$\$\_\{-10\}\$ &                                                NaN \\
            Laurentia-Scotland &                                     Torridon Group &      B &     354.3 &      57.9 & 220.9 & -17.7 &       7.1 &  925\$\textasciicircum \{+145\}\$\$\_\{-145\}\$ &                        Nordic workshop calculation \\
        Baltica - Fennoscandia &                Bjerkreim-Sokndal layered intrusion &      B &       6.1 &      58.5 & 217.9 & -35.9 &       6.3 &    921\$\textasciicircum \{+17\}\$\$\_\{-17\}\$ &                                                NaN \\
                       Baltica &                    Bratton and Algon Igneous Rocks &      B &      11.7 &      57.9 & 249.0 &   5.0 &       3.9 &    916\$\textasciicircum \{+11\}\$\$\_\{-11\}\$ &                                                NaN \\
                       Baltica &                           MEAN Rogaland igneous cx &      B &       6.9 &      58.3 & 207.9 & -43.2 &      10.1 &    903\$\textasciicircum \{+33\}\$\$\_\{-33\}\$ &                                                NaN \\
                       Baltica &                                   Hunnedalen dikes &      B &       7.0 &      58.9 & 222.0 & -41.0 &      15.6 &    848\$\textasciicircum \{+27\}\$\$\_\{-27\}\$ &                                                NaN \\
            Laurentia-Svalbard &                      Lower Grusdievbreen Formation &      B &      18.0 &      79.0 & 204.9 &  19.6 &      10.9 &    831\$\textasciicircum \{+20\}\$\$\_\{-20\}\$ &                                 \textbackslash cite\{Maloof2006a\} \\
            Laurentia-Svalbard &                      Upper Grusdievbreen Formation &      B &      18.2 &      78.9 & 252.6 &  -1.1 &       6.2 &    800\$\textasciicircum \{+11\}\$\$\_\{-11\}\$ &                                 \textbackslash cite\{Maloof2006a\} \\
                       Baltica &                                    Katav Formation &      B &      57.4 &      54.8 & 169.9 &  35.7 &      11.4 &  800\$\textasciicircum \{+100\}\$\$\_\{-100\}\$ &                                                NaN \\
                     Laurentia &                                    Gunbarrel dikes &      B &     248.7 &      44.8 & 138.2 &   9.1 &      12.0 &      778\$\textasciicircum \{+2\}\$\$\_\{-2\}\$ &  Calculation from \textbackslash cite\{Eyster2019a\} based on d... \\
            Laurentia-Svalbard &                          Svanbergfjellet Formation &      B &      18.0 &      78.5 & 226.8 &  25.9 &       5.8 &    770\$\textasciicircum \{+19\}\$\$\_\{-40\}\$ &                                 \textbackslash cite\{Maloof2006a\} \\
                     Laurentia &                               Uinta Mountain Group &      B &     250.7 &      40.8 & 161.3 &   0.8 &       4.7 &     760\$\textasciicircum \{+6\}\$\$\_\{-10\}\$ &                                   \textbackslash cite\{Weil2006b\} \\
                     Laurentia &                                      Carbon Canyon &     NR &     248.2 &      36.1 & 166.0 &  -0.5 &       9.7 &      757\$\textasciicircum \{+7\}\$\$\_\{-7\}\$ &  \textbackslash cite\{Weil2004a\} as calculated in \textbackslash cite\{Eyster... \\
                     Laurentia &                               Carbon Butte/Awatubi &     NR &     248.5 &      35.2 & 163.8 &  14.2 &       3.5 &      751\$\textasciicircum \{+8\}\$\$\_\{-8\}\$ &                                 \textbackslash cite\{Eyster2019a\} \\
                     Laurentia &                          Franklin event grand mean &      A &     275.4 &      73.0 & 162.1 &   6.7 &       3.0 &      718\$\textasciicircum \{+2\}\$\$\_\{-2\}\$ &                               \textbackslash cite\{Denyszyn2009b\} \\
                       Baltica &                                     Egersund dikes &      A &       5.9 &      58.8 &  44.1 &  31.4 &      15.6 &      616\$\textasciicircum \{+3\}\$\$\_\{-3\}\$ &                                                NaN \\
                     Laurentia &                                   Long Range dikes &      B &     303.3 &      53.7 & 175.3 & -19.0 &      17.4 &      615\$\textasciicircum \{+2\}\$\$\_\{-2\}\$ &                                 \textbackslash cite\{Murthy1992a\} \\
                     Laurentia &                           Baie des Moutons complex &      B &     301.0 &      50.8 & 152.7 & -42.6 &      12.0 &      583\$\textasciicircum \{+2\}\$\$\_\{-2\}\$ &                             \textbackslash cite\{McCausland2011a\} \\
                     Laurentia &                           Baie des Moutons complex &      B &     301.0 &      50.8 & 141.5 &  34.2 &      15.4 &      583\$\textasciicircum \{+2\}\$\$\_\{-2\}\$ &                             \textbackslash cite\{McCausland2011a\} \\
                     Laurentia &                         Callander Alkaline Complex &      B &     280.6 &      46.2 & 121.4 & -46.3 &       6.0 &      575\$\textasciicircum \{+5\}\$\$\_\{-5\}\$ &                                 \textbackslash cite\{Symons1991a\} \\
                     Laurentia &                                   Catoctin Basalts &      B &     281.8 &      38.5 & 116.7 & -42.0 &      17.5 &      572\$\textasciicircum \{+5\}\$\$\_\{-5\}\$ &                                  \textbackslash cite\{Meert1994a\} \\
                       Baltica &                                       Volhyn lavas &      B &      26.0 &      51.2 &   4.4 &  19.8 &      28.3 &    571\$\textasciicircum \{+10\}\$\$\_\{-10\}\$ &                                                NaN \\
                     Laurentia &                        Sept-Iles layered intrusion &      B &     293.5 &      50.2 & 141.0 &  20.0 &       6.7 &      565\$\textasciicircum \{+4\}\$\$\_\{-4\}\$ &                                \textbackslash cite\{Tanczyk1987a\} \\
                       Baltica &                               Kurgashlya formation &      A &      57.5 &      53.3 & 314.5 &  50.9 &       5.3 &      565\$\textasciicircum \{+5\}\$\$\_\{-5\}\$ &                                                NaN \\
                       Baltica &                                  Bakeevo formation &      A &      58.2 &      54.9 & 299.1 &  42.3 &       5.3 &      565\$\textasciicircum \{+5\}\$\$\_\{-5\}\$ &                                                NaN \\
                       Baltica &                        Basu-Kukkarauk formation -C &      B &      57.0 &      54.0 & 187.3 &   1.1 &       5.8 &      562\$\textasciicircum \{+5\}\$\$\_\{-5\}\$ &                                                NaN \\
                 Baltica-Urals &                          Basu formation  (precise) &      B &      56.9 &      53.9 & 186.1 &   1.7 &       3.8 &    560\$\textasciicircum \{+10\}\$\$\_\{-10\}\$ &                                                NaN \\
                       Baltica &                 Chernokamenskay group sediments -C &      B &      58.4 &      58.5 & 306.7 &  17.3 &       6.0 &    557\$\textasciicircum \{+13\}\$\$\_\{-13\}\$ &                                                NaN \\
                       Baltica &                             MEAN Vendian sediments &      B &      40.0 &      65.5 & 293.0 &  31.3 &       9.9 &      556\$\textasciicircum \{+1\}\$\$\_\{-1\}\$ &                                                NaN \\
                 Baltica-Urals &                                Zigan HTC thin beds &      B &      56.7 &      54.4 & 138.3 & -18.0 &      10.7 &   548\$\textasciicircum \{+27\}\$\$\_\{--12\}\$ &                                                NaN \\
                       Baltica &                   Zigan formation clastic rocks -C &      B &      56.7 &      53.7 & 138.4 & -16.2 &       4.1 &      548\$\textasciicircum \{+4\}\$\$\_\{-4\}\$ &                                                NaN \\
\end{longtable}