### 2020-09-02 Manuscript/pole_table_edited.tex

\begin{longtable}{p{1 in}p{1 in}rrrrrrrp{1.5 in}}
\toprule
                       terrane &                                        unit name & Nordic rating &  site lon (\textdegree) &  site lat (\textdegree) &  plon (\textdegree) &  plat (\textdegree) &  A$_{95}$ (\textdegree) &                    age (Ma)&                                     pole reference \\ \hline
\midrule
\endhead
\midrule
\multicolumn{10}{r}{{Continued on next page}} \\ \hline
\midrule
\endfoot

\bottomrule
\endlastfoot
             Laurentia-Wyoming &                          Stillwater Complex - C2 &      A &     249.2 &      45.2 & 335.8 & -83.6 &       4.0 &     2705$^{+4}_{-4}$ &                                 \cite{Selkin2008a} \\ \hline
      Laurentia-Superior(East) &                     Otto Stock dykes and aureole &      B &     279.9 &      48.0 & 227.0 &  69.0 &       4.8 &     2676$^{+5}_{-5}$ &                               \cite{Pullaiah1975b} \\ \hline
               Laurentia-Slave &                                     Defeat Suite &      B &     245.5 &      62.5 &  64.0 &  -1.0 &      15.0 &     2625$^{+5}_{-5}$ &                               \cite{Mitchell2014a} \\ \hline
      Laurentia-Superior(East) &                       Ptarmigan-Mistassini dykes &      B &     287.0 &      54.0 & 213.0 & -45.3 &      13.8 &     2505$^{+2}_{-2}$ &                                  \cite{Evans2010a} \\ \hline
      Laurentia-Superior(East) &                               Matachewan dykes R &      A &     278.0 &      48.0 & 238.3 & -44.1 &       1.6 &   2466$^{+23}_{-23}$ &                                  \cite{Evans2010a} \\ \hline
      Laurentia-Superior(East) &                               Matachewan dykes N &      A &     278.0 &      48.0 & 239.5 & -52.3 &       2.4 &     2446$^{+3}_{-3}$ &                                  \cite{Evans2010a} \\ \hline
               Laurentia-Slave &                                     Malley dykes &      A &     249.8 &      64.2 & 310.0 & -50.8 &       6.7 &     2231$^{+2}_{-2}$ &                                 \cite{Buchan2012a} \\ \hline
      Laurentia-Superior(East) &                                 Senneterre dykes &      A &     283.0 &      49.0 & 284.3 & -15.3 &       5.5 &     2218$^{+6}_{-6}$ &                                 \cite{Buchan1993a} \\ \hline
      Laurentia-Superior(East) &                               Nipissing N1 sills &      A &     279.0 &      47.0 & 272.0 & -17.0 &      10.0 &     2217$^{+4}_{-4}$ &                                 \cite{Buchan2000a} \\ \hline
               Laurentia-Slave &                                     Dogrib dykes &      A &     245.5 &      62.5 & 315.0 & -31.0 &       7.0 &     2193$^{+2}_{-2}$ &                               \cite{Mitchell2014a} \\ \hline
      Laurentia-Superior(East) &                                Biscotasing dykes &      A &     280.0 &      48.0 & 223.9 &  26.0 &       7.0 &     2170$^{+3}_{-3}$ &                                  \cite{Evans2010a} \\ \hline
             Laurentia-Wyoming &  Rabbit Creek, Powder River and South Path Dykes &      A &     252.8 &      43.9 & 339.2 &  65.5 &       7.6 &    2160$^{+11}_{-8}$ &                                 \cite{Kilian2015a} \\ \hline
               Laurentia-Slave &                                      Indin dykes &      A &     245.6 &      62.5 & 256.0 & -36.0 &       7.0 &    2126$^{+3}_{-18}$ &                                 \cite{Buchan2016a} \\ \hline
      Laurentia-Superior(West) &                                 Marathon dykes N &      A &     275.0 &      49.0 & 198.2 &  45.4 &       7.7 &     2124$^{+3}_{-3}$ &                                  \cite{Halls2008a} \\ \hline
      Laurentia-Superior(West) &                                 Marathon dykes R &      A &     275.0 &      49.0 & 182.2 &  55.1 &       7.5 &     2104$^{+3}_{-3}$ &                                  \cite{Halls2008a} \\ \hline
      Laurentia-Superior(West) &                               Cauchon Lake dykes &      A &     263.0 &      56.0 & 180.9 &  53.8 &       7.7 &     2091$^{+2}_{-2}$ &                                  \cite{Evans2010a} \\ \hline
      Laurentia-Superior(West) &                               Fort Frances dykes &      A &     266.0 &      48.0 & 184.6 &  42.8 &       6.1 &     2077$^{+5}_{-5}$ &                                  \cite{Evans2010a} \\ \hline
      Laurentia-Superior(East) &                                 Lac Esprit dykes &      A &     282.0 &      53.0 & 170.5 &  62.0 &       6.4 &     2069$^{+1}_{-1}$ &                                  \cite{Evans2010a} \\ \hline
      Laurentia-Greenland-Nain &                                  Kangamiut Dykes &      B &     307.0 &      66.0 & 273.8 &  17.1 &       2.7 &   2042$^{+12}_{-12}$ &                                 \cite{Fahrig1976b} \\ \hline
               Laurentia-Slave &                                Lac de Gras dykes &      A &     249.6 &      64.4 & 267.9 &  11.8 &       7.1 &     2026$^{+5}_{-5}$ &                                 \cite{Buchan2009a} \\ \hline
      Laurentia-Superior(East) &                                      Minto dykes &      A &     285.0 &      57.0 & 171.5 &  38.7 &      13.1 &     1998$^{+2}_{-2}$ &                                  \cite{Evans2010a} \\ \hline
               Laurentia-Slave &                                  Rifle Formation &      B &     252.9 &      65.9 & 341.0 &  14.0 &       7.7 &     1963$^{+6}_{-6}$ &                                  \cite{Evans1981a} \\ \hline
                 Laurentia-Rae &                           Clearwater Anorthosite &      B &     251.6 &      57.1 & 311.8 &   6.5 &       2.9 &     1917$^{+7}_{-7}$ &                                  \cite{Halls1999a} \\ \hline
             Laurentia-Wyoming &                       Sourdough mafic dike swarm &      A &    -108.3 &      44.7 & 292.0 &  49.2 &       8.1 &     1899$^{+5}_{-5}$ &                                 \cite{Kilian2016b} \\ \hline
               Laurentia-Slave &                                 Ghost Dike Swarm &      A &     244.6 &      62.6 & 286.0 &  -2.0 &       6.0 &     1887$^{+5}_{-9}$ &                                 \cite{Buchan2016a} \\ \hline
               Laurentia-Slave &                         Mean Seton/Akaitcho/Mara &      B &     250.0 &      65.0 & 260.0 &  -6.0 &       4.0 &     1885$^{+5}_{-5}$ &                               \cite{Mitchell2010c} \\ \hline
               Laurentia-Slave &                   Mean Kahochella, Peacock Hills &      B &     250.0 &      65.0 & 285.0 & -12.0 &       7.0 &     1882$^{+4}_{-4}$ &                               \cite{Mitchell2010c} \\ \hline
      Laurentia-Superior(West) &                              Molson (B+C2) dykes &      A &     262.0 &      55.0 & 218.0 &  28.9 &       3.8 &     1879$^{+6}_{-6}$ &                                  \cite{Evans2010a} \\ \hline
               Laurentia-Slave &        Douglas Peninsula Formation, Pethei Group &      B &     249.7 &      62.8 & 258.0 & -18.0 &      14.2 &   1876$^{+10}_{-10}$ &                                \cite{Irving1979a} \\ \hline
               Laurentia-Slave &                               Takiyuak Formation &      B &     246.9 &      66.1 & 249.0 & -13.0 &       8.0 &   1876$^{+10}_{-10}$ &                                \cite{Irving1979a} \\ \hline
            Laurentia-Superior &                       Haig/Flaherty/Sutton Mean  &      B &     279.0 &      56.0 & 245.8 &   1.0 &       3.9 &     1870$^{+1}_{-1}$ &  Nordic workshop calculation based on data of \cite{Schmidt1980a, Schwarz1982a} \\ \hline
               Laurentia-Slave &             Pearson A/Peninsular/Kilohigok sills &      A &     250.0 &      65.0 & 269.0 & -22.0 &       6.0 &     1870$^{+4}_{-4}$ &                               \cite{Mitchell2010c} \\ \hline
 Laurentia-Trans-Hudson orogen &                              Boot-Phantom Pluton &      B &     258.1 &      54.7 & 275.4 &  62.4 &       7.9 &     1838$^{+1}_{-1}$ &                                 \cite{Symons1999a} \\ \hline
                 Laurentia-Rae &                                    Sparrow Dykes &      B &     250.2 &      61.6 & 291.0 &  12.0 &       7.9 &     1827$^{+4}_{-4}$ &                                \cite{McGlynn1974a} \\ \hline
                 Laurentia-Rae &                                 Martin Formation &      A &     251.4 &      59.6 & 288.0 &  -9.0 &       8.5 &     1818$^{+4}_{-4}$ &                                  \cite{Evans1973a} \\ \hline
                     Laurentia &                                    Dubawnt Group &      B &     265.6 &      64.1 & 277.0 &   7.0 &       8.0 &   1785$^{+35}_{-35}$ &                                    \cite{Park1973a} \\ \hline
 Laurentia-Trans-Hudson orogen &                          Deschambault Pegmatites &      B &     256.7 &      54.9 & 276.0 &  67.5 &       7.7 &     1766$^{+5}_{-5}$ &                                 \cite{Symons2000a} \\ \hline
 Laurentia-Trans-Hudson orogen &                                 Jan Lake Granite &      B &     257.2 &      54.9 & 264.3 &  24.3 &      16.9 &     1758$^{+1}_{-1}$ &                                   \cite{Gala1995a} \\ \hline
                     Laurentia &                                    Cleaver Dykes &      A &     242.0 &      67.5 & 276.7 &  19.4 &       6.1 &     1741$^{+5}_{-5}$ &                                  \cite{Irving2004a} \\ \hline
           Laurentia-Greenland &                      Melville Bugt diabase dykes &      B &     303.0 &      74.6 & 273.8 &   5.0 &       8.7 &     1633$^{+5}_{-5}$ &                                  \cite{Halls2011a} \\ \hline
                     Laurentia &                          Western Channel Diabase &      A &     242.2 &      66.4 & 245.0 &   9.0 &       6.6 &     1590$^{+3}_{-3}$ &                                 \cite{Irving1972a} \\ \hline
                     Laurentia &               St.Francois Mountains Acidic Rocks &      A &     269.5 &      37.5 & 219.0 & -13.2 &       6.1 &   1476$^{+16}_{-16}$ &                                  \cite{Meert2002b} \\ \hline
                     Laurentia &                             Michikamau Intrusion &      A &     296.0 &      54.5 & 217.5 &  -1.5 &       4.7 &     1460$^{+5}_{-5}$ &                                 \cite{Emslie1976a} \\ \hline
                     Laurentia &                                Spokane Formation &      A &     246.8 &      48.2 & 215.5 & -24.8 &       4.7 &   1458$^{+13}_{-13}$ &                                 \cite{Elston2002a} \\ \hline
                     Laurentia &                               Snowslip Formation &      A &     245.9 &      47.9 & 210.2 & -24.9 &       3.5 &   1450$^{+14}_{-14}$ &                                 \cite{Elston2002a} \\ \hline
                     Laurentia &                               Tobacco Root dykes &      B &     247.6 &      47.4 & 216.1 &   8.7 &      10.5 &   1448$^{+49}_{-49}$ &                                 \cite{Harlan2008a} \\ \hline
                     Laurentia &                                     Purcell Lava &      A &     245.1 &      49.4 & 215.6 & -23.6 &       4.8 &     1443$^{+7}_{-7}$ &                                 \cite{Elston2002a} \\ \hline
                     Laurentia &                        Rocky Mountain intrusions &      B &     253.8 &      40.3 & 217.4 & -11.9 &       9.7 &   1430$^{+15}_{-15}$ &  Nordic workshop calculation based on data of \cite{Harlan1994a,Harlan1998a} \\ \hline
                     Laurentia &                                 Mistastin Pluton &      B &     296.3 &      55.6 & 201.5 &  -1.0 &       7.6 &   1425$^{+25}_{-25}$ &                                 \cite{Fahrig1976a} \\ \hline
                     Laurentia &                               McNamara Formation &      A &     246.4 &      46.9 & 208.3 & -13.5 &       6.7 &     1401$^{+6}_{-6}$ &                                 \cite{Elston2002a} \\ \hline
                     Laurentia &       Pilcher, Garnet Range and Libby Formations &      A &     246.4 &      46.7 & 215.3 & -19.2 &       7.7 &   1385$^{+23}_{-23}$ &                                 \cite{Elston2002a} \\ \hline
           Laurentia-Greenland &                              Zig-Zag Dal Basalts &      B &     334.8 &      81.2 & 242.8 &  12.0 &       3.8 &     1382$^{+2}_{-2}$ &                              \cite{Marcussen1983a} \\ \hline
           Laurentia-Greenland &                            Midsommersoe Dolerite &      B &     333.4 &      81.6 & 242.0 &   6.9 &       5.1 &     1382$^{+2}_{-2}$ &                              \cite{Marcussen1983a} \\ \hline
           Laurentia-Greenland &                    Victoria Fjord dolerite dykes &      B &     315.3 &      81.5 & 231.7 &  10.3 &       4.3 &     1382$^{+2}_{-2}$ &                             \cite{Abrahamsen1987a} \\ \hline
                     Laurentia &                                 Nain Anorthosite &      B &     298.2 &      56.5 & 206.7 &  11.7 &       2.2 &   1305$^{+15}_{-15}$ &                                \cite{Murthy1978a} \\ \hline
           Laurentia-Greenland &                           North Qoroq intrusives &      B &     314.6 &      61.1 & 202.6 &  13.2 &       8.3 &     1275$^{+1}_{-1}$ &                                  \cite{Piper1992a} \\ \hline
           Laurentia-Greenland &                                Kungnat Ring Dyke &      B &     311.7 &      61.2 & 198.7 &   3.4 &       3.2 &     1275$^{+2}_{-2}$ &                                  \cite{Piper1977b} \\ \hline
                     Laurentia &                       Mackenzie dykes grand mean &      A &     250.0 &      65.0 & 190.0 &   4.0 &       5.0 &     1267$^{+2}_{-2}$ &                                 \cite{Buchan2000a} \\ \hline
           Laurentia-Greenland &                       West Gardar Dolerite Dykes &      B &     311.7 &      61.2 & 201.7 &   8.7 &       6.6 &     1244$^{+8}_{-8}$ &                                  \cite{Piper1977b} \\ \hline
           Laurentia-Greenland &                    West Gardar Lamprophyre Dykes &      B &     311.7 &      61.2 & 206.4 &   3.2 &       7.2 &   1238$^{+11}_{-11}$ &                                  \cite{Piper1977b} \\ \hline
                     Laurentia &                           Sudbury Dykes Combined &      A &     278.6 &      46.3 & 192.8 &  -2.5 &       2.5 &     1237$^{+5}_{-5}$ &                                 \cite{Palmer1977a} \\ \hline
            Laurentia-Scotland &                                      Stoer Group &      B &     354.5 &      58.0 & 238.4 &  37.2 &       7.7 &   1199$^{+70}_{-70}$ &                        Nordic workshop calculation \\ \hline
           Laurentia-Greenland &                                   Narssaq Gabbro &      B &     313.8 &      60.9 & 225.4 &  31.6 &       9.7 &     1184$^{+5}_{-5}$ &                                  \cite{Piper1977a} \\ \hline
           Laurentia-Greenland &                               Hviddal Giant Dyke &      B &     313.7 &      60.9 & 215.3 &  33.2 &       9.6 &     1184$^{+5}_{-5}$ &                                  \cite{Piper1977a} \\ \hline
           Laurentia-Greenland &                                South Qoroq Intr. &      A &     314.6 &      61.1 & 215.9 &  41.8 &      13.1 &     1163$^{+2}_{-2}$ &                                  \cite{Piper1992a} \\ \hline
           Laurentia-Greenland &                               Giant Gabbro Dykes &      B &     313.7 &      60.9 & 226.1 &  42.3 &       9.4 &     1163$^{+2}_{-2}$ &                                  \cite{Piper1977a} \\ \hline
           Laurentia-Greenland &                             NE-SW Trending dykes &      B &     314.6 &      61.1 & 230.8 &  33.4 &       5.7 &     1160$^{+5}_{-5}$ &                                  \cite{Piper1992a} \\ \hline
                     Laurentia &                        Ontario lamprophyre dykes &      NR &     273.3 &      48.8 & 223.3 &  58.0 &       9.2 &   1143$^{+10}_{-10}$ &                                 \cite{Piispa2018a} \\ \hline
                     Laurentia &                                    Abitibi Dykes &      A &     279.0 &      48.0 & 215.5 &  48.8 &      14.1 &     1141$^{+2}_{-2}$ &                                  \cite{Ernst1993a} \\ \hline
                     Laurentia &                          Nipigon sills and lavas &      A &     270.9 &      49.1 & 217.8 &  47.2 &       4.0 &     1109$^{+2}_{-2}$ &  Nordic workshop calculation based on data of \cite{Palmer1970a, Robertson1971a, Pesonen1979a, Pesonen1979b, Middleton2004a, Borradaile2006a} \\ \hline
                     Laurentia &           Lowermost Mamainse Point volcanics -R1 &      A &     275.3 &      47.1 & 227.0 &  49.5 &       5.3 &     1109$^{+2}_{-3}$ &                         \cite{Swanson-Hysell2014a} \\ \hline
                     Laurentia &                         Lower Osler volcanics -R &      A &     272.3 &      48.8 & 218.6 &  40.9 &       4.8 &     1108$^{+3}_{-3}$ &                         \cite{Swanson-Hysell2014b} \\ \hline
                     Laurentia &                        Middle Osler volcanics -R &      A &     272.4 &      48.8 & 211.3 &  42.7 &       8.2 &     1107$^{+4}_{-4}$ &                         \cite{Swanson-Hysell2014b} \\ \hline
                     Laurentia &                         Upper Osler volcanics -R &      A &     272.4 &      48.7 & 203.4 &  42.3 &       3.7 &     1105$^{+1}_{-1}$ &  \cite{Halls1974a, Swanson-Hysell2014b, Swanson-Hysell2019a} \\ \hline
                     Laurentia &               Lower Mamainse Point volcanics -R2 &      A &     275.3 &      47.1 & 205.2 &  37.5 &       4.5 &     1105$^{+3}_{-4}$ &                         \cite{Swanson-Hysell2014a} \\ \hline
                     Laurentia &   Mamainse Point volcanics -C (lower N, upper R) &      A &     275.3 &      47.1 & 189.7 &  36.1 &       4.9 &     1101$^{+1}_{-1}$ &                         \cite{Swanson-Hysell2014a} \\ \hline
                     Laurentia &                             North Shore lavas -N &      A &     268.7 &      46.3 & 181.7 &  31.1 &       2.1 &     1097$^{+3}_{-3}$ &              \cite{Tauxe2009a,Swanson-Hysell2019a} \\ \hline
                     Laurentia &                           Portage Lake Volcanics &      A &     271.2 &      47.0 & 182.5 &  27.5 &       2.3 &     1095$^{+3}_{-3}$ &  \cite{Books1972a, Hnat2006a} as calculated in \cite{Swanson-Hysell2019a} \\ \hline
                     Laurentia &                            Chengwatana Volcanics &      B &     267.3 &      45.4 & 186.1 &  30.9 &       8.2 &     1095$^{+2}_{-2}$ &                                  \cite{Kean1997a} \\ \hline
                     Laurentia &            Uppermost Mamainse Point volcanics -N &      A &     275.3 &      47.1 & 183.2 &  31.2 &       2.5 &     1094$^{+6}_{-4}$ &                         \cite{Swanson-Hysell2014a} \\ \hline
                     Laurentia &                  Cardenas Basalts and Intrusions &      B &     248.1 &      36.1 & 185.0 &  32.0 &       8.0 &     1091$^{+5}_{-5}$ &                                   \cite{Weil2003a} \\ \hline
                     Laurentia &                         Schroeder Lutsen Basalts &      A &     269.1 &      47.5 & 187.8 &  27.1 &       3.0 &     1090$^{+2}_{-7}$ &                              \cite{Fairchild2017a} \\ \hline
                     Laurentia &                      Central Arizona diabases -N &      A &     249.2 &      33.7 & 175.3 &  15.7 &       7.0 &   1088$^{+11}_{-11}$ &                               \cite{Donadini2011b} \\ \hline
                     Laurentia &                                 Lake Shore Traps &      A &     271.9 &      47.6 & 186.4 &  23.1 &       4.0 &     1086$^{+1}_{-1}$ &                                \cite{Kulakov2013a} \\ \hline
                     Laurentia &                    Michipicoten Island Formation &      A &     274.3 &      47.7 & 174.7 &  17.0 &       4.4 &     1084$^{+1}_{-1}$ &                              \cite{Fairchild2017a} \\ \hline
                     Laurentia &                                   Nonesuch Shale &      B &     271.5 &      47.0 & 178.1 &   7.6 &       5.5 &    1080$^{+4}_{-10}$ &                                  \cite{Henry1977a} \\ \hline
                     Laurentia &                                  Freda Sandstone &      B &     271.5 &      47.0 & 179.0 &   2.2 &       4.2 &   1070$^{+14}_{-10}$ &                                  \cite{Henry1977a} \\ \hline
                     Laurentia &                            Haliburton Intrusions &      B &     281.4 &      45.0 & 141.9 & -32.6 &       6.3 &   1015$^{+15}_{-15}$ &                                \cite{Warnock2000a} \\ \hline
            Laurentia-Scotland &                                   Torridon Group &      B &     354.3 &      57.9 & 220.9 & -17.7 &       7.1 &  925$^{+145}_{-145}$ &                        Nordic workshop calculation \\ \hline
            Laurentia-Svalbard &                    Lower Grusdievbreen Formation &      B &      18.0 &      79.0 & 204.9 &  19.6 &      10.9 &    831$^{+20}_{-20}$ &                                 \cite{Maloof2006a} \\ \hline
            Laurentia-Svalbard &                    Upper Grusdievbreen Formation &      B &      18.2 &      78.9 & 252.6 &  -1.1 &       6.2 &    800$^{+11}_{-11}$ &                                 \cite{Maloof2006a} \\ \hline
            Laurentia &                                  Gunbarrel dykes &      B &     248.7 &      44.8 & 138.2 &   9.1 &      12.0 &      778$^{+2}_{-2}$ &  Calculation from \cite{Eyster2019a} based on data of \cite{Harlan1993a, Harlan1997a}  \\ \hline
   Laurentia-Svalbard &                        Svanbergfjellet Formation &      B &      18.0 &      78.5 & 226.8 &  25.9 &       5.8 &    770$^{+19}_{-40}$ &                                 \cite{Maloof2006a} \\ \hline
            Laurentia &                             Uinta Mountain Group &      B &     250.7 &      40.8 & 161.3 &   0.8 &       4.7 &     760$^{+6}_{-10}$ &                                   \cite{Weil2006b} \\ \hline
            Laurentia &                                    Carbon Canyon &     NR &     248.2 &      36.1 & 166.0 &  -0.5 &       9.7 &      757$^{+7}_{-7}$ &  \cite{Weil2004a} as calculated in \cite{Eyster2019a} \\ \hline
            Laurentia &                             Carbon Butte/Awatubi &     NR &     248.5 &      35.2 & 163.8 &  14.2 &       3.5 &      751$^{+8}_{-8}$ &                                 \cite{Eyster2019a} \\ \hline
            Laurentia &                        Franklin event grand mean &      A &     275.4 &      73.0 & 162.1 &   6.7 &       3.0 &      724$^{+3}_{-3}$ &                               \cite{Denyszyn2009a} \\ \hline
                     Laurentia &                                 Long Range Dykes &      B &     303.3 &      53.7 & 175.3 & -19.0 &      17.4 &      615$^{+2}_{-2}$ &                                 \cite{Murthy1992a} \\ \hline
                     Laurentia &                         Baie des Moutons complex (A) &      B &     301.0 &      50.8 & 152.7 & -42.6 &      12.0 &      583$^{+2}_{-2}$ &                             \cite{McCausland2011a} \\ \hline
                     Laurentia &                         Baie des Moutons complex (B) &      B &     301.0 &      50.8 & 141.5 &  34.2 &      15.4 &      583$^{+2}_{-2}$ &                             \cite{McCausland2011a} \\ \hline
                     Laurentia &                       Callander Alkaline Complex &      B &     280.6 &      46.2 & 121.4 & -46.3 &       6.0 &      575$^{+5}_{-5}$ &                                 \cite{Symons1991a} \\ \hline
                     Laurentia &                                 Catoctin Basalts &      B &     281.8 &      38.5 & 116.7 & -42.0 &      17.5 &      572$^{+5}_{-5}$ &                                  \cite{Meert1994a} \\ \hline
                     Laurentia &                      Sept-Iles layered intrusion &      B &     293.5 &      50.2 & 141.0 &  20.0 &       6.7 &      565$^{+4}_{-4}$ &                                 \cite{Tanczyk1987a} \\ \hline
\end{longtable}